---
title: "CMTH 642 - Assignment 2"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# USDA Clean Data

We uplodaded the clean csv file generated from Assignment 1 (USDA_Clean.csv).
Please download and load it to your workspace.

```{r}
USDAclean = read.csv("USDA_Clean.csv")
#attach(USDA_Clean) ## Optional
# attch() function helps you to access USDA_Clean without the need of menioning it.
# For example, you can use Calories instead of USDA_Clean$Calories
#View(USDA_Clean)
# str(USDAclean)
```


# Visualization of Feature Relationships

We have used a function panel.cor() inside pair() to show the correlations among different features. The only line you should complete is the line that you assign a value to **USDA_Selected_Featuers**. Research how can you select multiple columns from a dataframe to use it inside pair() function. 

A) Show the relationship among *Calories*, *Carbohydrate*, *Protein*, *Total Fat* and *Sodium*. **(5 p)**

B) Describe the correlations among **Calories**  and other features. **(5 p)**

Hint: We usually interpret the absolute value of correlation as follows:

.00-.19 *very weak*

.20-.39 *weak*

.40-.59 *moderate*

.60-.79 *strong*

.80-1.0 *very strong*


```{r}
panel.cor <- function(x, y, digits=2, prefix="", cex.cor) 
{
    usr <- par("usr"); on.exit(par(usr)) 
    par(usr = c(0, 1, 0, 1)) 
    r <- abs(cor(x, y)) 
    txt <- format(c(r, 0.123456789), digits=digits)[1] 
    txt <- paste(prefix, txt, sep="") 
    if(missing(cex.cor)) cex <- 0.8/strwidth(txt) 
 
    test <- cor.test(x,y) 
    # borrowed from printCoefmat
    Signif <- symnum(test$p.value, corr = FALSE, na = FALSE, 
                  cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
                  symbols = c("***", "**", "*", ".", " ")) 
 
    text(0.5, 0.5, txt, cex = cex * r) 
    text(.8, .8, Signif, cex=cex, col=2) 
}
# Assign a value USDA_Selected_Featuers that represents
# "Calories","Carbohydrate","Protein","TotalFat", "Sodium" columns
###################################################
##### Complete code here and uncomment it
USDA_Selected_Featuers <- USDAclean[,c('Calories','Carbohydrate','Protein','TotalFat','Sodium')]
###################################################

#### Uncomment the following line when you assign USDA_Selected_Featuers to show the results
pairs(USDA_Selected_Featuers, lower.panel=panel.smooth, upper.panel=panel.cor)

# Explain what you can conclude from this visualization as a comment here
# Calories and TotalFat have the highest correlation: (0.8). Second is Calroies and Carbohydrate (0.42).

```

## Regression Model on USDA Clean Data
Create a Linear Regression Model (lm), using **Calories** as the dependent variable, and *Carbohydrate*, *Protein*, *Total Fat* and *Sodium* as independent variables. **(10 p)**
```{r}
# Write your code here
food.lm <- lm(Calories ~ Carbohydrate + Protein + TotalFat + Sodium, data=USDA_Selected_Featuers)

for(attr in c('Carbohydrate','Protein','TotalFat','Sodium')){
  USDA_Selected_Featuers[,attr] <- as.factor(USDA_Selected_Featuers[,attr])
}

# Below is a shorthand (since we have already subsetted the dataframe): 
# food.lm <- lm(Calories ~ ., data=USDA_Selected_Featuers)
```

## Analyzing Regression Model
A) In the above example, which independent feature is less significant? (Hint: Use ANOVA) **(5 p)**


```{r}
# Write your code here and explain answers as a comment
anova(food.lm)
# Sodium is least significant (has highest p-value)
```
B) Which independent variable has the strongest positive predictive power in the model? (Hint: Look at the coefitients calculated for each independant variable) **(5 p)**
```{r}
# Write your code here and explain answers as a comment
summary(food.lm)
# I'm guessing that TotalFat has the highest predictive power because it has the largest co-efficient
```
## Calories Prediction
A new product is just produced with the following data:

"Protein"	"TotalFat"	"Carbohydrate"	"Sodium"	"Cholesterol"

0.1       40        425             430       75 

"Sugar"	"Calcium"	"Iron"	"Potassium"	"VitaminC"	"VitaminE"	"VitaminD"

NA      42        NA      35			    10			    0.0			    NA

A) Based on the model you created, what is the predicted value for **Calories** ? **(5 p)**
```{r}
predict(food.lm, data.frame(Protein=0.1,TotalFat=40,Carbohydrate=425,Sodium=430,Cholesterol=75),interval="prediction")
# Predicted Value: 1943.65
```

B) If the *Sodium* amount increases 101 times from 430 to 43430 (10000% increase), how much change will occur on Calories in percent? Can you explain why? **(5 p)**

```{r}
# Write your code here and explain answers as a comment

predict(food.lm, data.frame(Protein=0.1,TotalFat=40,Carbohydrate=425,Sodium=43430,Cholesterol=75),interval="prediction")
# Increasing the sodium from 430 to 43430 increases the predicted calories from 1943.65 to 1957.622
# This is because Sodium has the lowest coefficient in the model (0.0003 vs 3.736 for the next lowest, Carbohydrate)
```

# Wilcoxon Tests

## Research Question: Does illustrations improve memorization? 
A study of primary education asked elementaty school students to retell two book articles that they read earlier in the week. The first (Article 1) had no picutres, and the second (Article 2) illustrated with pictures. An expert listened
to recordings of the students retelling each article and assigned a score for
certain uses of language. Higher scores are better. Here are the data for five readers in a this study:

Student     1     2     3     4     5

Article 1   0.40  0.72  0.00  0.36  0.55

Article 2   0.77  0.49  0.66  0.28  0.38

We wonder if illustrations improve how the students retell an article. 

### What is $H_0$ and $H_a$ ? 

**(10 p)**

```{r}
# Write your Answer as a comment here

```

### Paired or Independent design?

Based on your answer, which Wilcoxon test should you use? **(5 p)**
```{r}
# Write your Answer as a comment here

```

### Will you accept or reject your Null Hypothesis? ($\alpha = 0.05$)
Do illustrations improve how the students retell an article or not? **(5 p)**

```{r}
# Write your code here

```

## Packaging Problem

Two companies selling toothpastes with the lable of 100 grams per tube on the package. We randomly bought eight toothpastes from each company A and B from random stores. Afterwards, we scaled them using high precision scale. Our measurements are recorded as follows:

Company A: 97.1   101.3   107.8   101.9   97.4    104.5   99.5    95.1

Company B: 103.5  105.3   106.5   107.9   102.1   105.6   109.8   97.2

### Distribution Analysis
Are the distributions of package weights similar for these companies? Are they normally distributed or skewed? **(10 p)** (Hint: Use boxplot)

```{r}
# Write your code here

```

### Are packaging process similar or different based on weight measurements? 

Can we be at least 95% confident that there is no difference between packaging of these two companies? **(5 p)** 

Can we be at least 99% confident? **(5 p)** 

Please explain.

```{r}
# Write your code here and explain answers as a comment

```
 
# Correlation
Plot and see the relationship between "cylinder" (cyl) and "weight" (wt) of the cars from mtcars dataset. 
A) Can you see any patterns of correlation between these two variable? **(5 p)**

```{r}
# Write your code here and explain answers as a comment

```

B) What is the best description for "cyl" and "wt" variables? (Ratio, Ordinal, Interval, or Categorical) **(5 p)**

```{r}
# Explain answers as a comment here

```

C) Based on the description of the "cyl" and "wt" variables, should you use "Pearson" or "Spearman" correlation? Find the correlation between these two variables. **(10 p)**

```{r}
# Write your code here and explain answers as a comment

```
