---
title: "CMTH 642, Lab 1"
output: pdf_document
---

1. Read the "train.csv" file from the following website.
"https://raw.githubusercontent.com/agconti/kaggle-titanic/master/data/train.csv"
```{r}
train.df <- read.csv("https://raw.githubusercontent.com/agconti/kaggle-titanic/master/data/train.csv")
```

2. Have a look at the data set.

3. Change the "Pclass" and "Survived" attributes to factors.
```{r}
train.df$Pclass <- as.factor(train.df$Pclass)
train.df$Survived <- as.factor(train.df$Survived)
```

4. Check the missing values of the "Age" and "Name" attributes. 
```{r}
train.df$Age[is.na(train.df$Age)]
train.df$Name[is.na(train.df$Name)]
```

5. For a title containing a missing value, assign the mean age value for each title not containing a missing value. 
(Hint: grepl(" Mr.\\.", train.data$Name))
```{r}
train.df[which(is.na(train.df$Age) & grepl("(Miss.|Mrs.)", train.df$Name)), "Age"] <- mean(train.df[which(!is.na(train.df$Age) & grepl("(Miss.|Mrs.)", train.df$Name)), "Age"], na.rm=T) #Women

train.df[which(is.na(train.df$Age) & grepl("(Master\\.|Mr\\.)", train.df$Name)), "Age"] <- mean(train.df[which(!is.na(train.df$Age) & grepl(" Mr.\\.", train.df$Name)), "Age"], na.rm=T) #Men

train.df[which(is.na(train.df$Age) & grepl(" Rev\\.", train.df$Name)), "Age"] <- mean(train.df[which(!is.na(train.df$Age) & grepl(" Rev.\\.", train.df$Name)), "Age"], na.rm=T) #Rev

train.df[which(is.na(train.df$Age) & grepl(" Dr\\.", train.df$Name)), "Age"] <- mean(train.df[which(!is.na(train.df$Age) & grepl(" Dr\\.", train.df$Name)), "Age"], na.rm=T) #Doctor
```

6. List the distribution of Port of Embarkation. Replace empty entries with NA for "Embarked" attribute.
```{r}
table(train.df$Embarked)
train.df[nchar(as.character(train.df$Embarked)) < 1,"Embarked"] <- NA
```

7. Assign the two missing values to the most counted port, which is Southampton in this case.

```{r}
train.df[is.na(train.df$Embarked),"Embarked"] <- "S" 
```

