---
title: "CMTH642 Lab 05"
author: "Geoffrey Clark"
date: "February 13, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Q.1: 
You want to see if the mean of goals suffered by two football teams over the years is the same.  T he number of goals suffered by each team in 6 games for each year are as follows:

Team X : 6, 8, 2, 4, 4, 5
Team Y : 7, 10, 4, 3, 5, 6
( ?? = 0.05 )

( Hint:  Wilcoxon rank sum test is used when is asked to compare the means of two groups that  do not follow a normal distribution: it is a non - parametrical test. It is the equivalent of the t test,  applied for independent samples)

```{r}
teamX <- c(6,8,2,4,4,5)
teamY <- c(7,10,4,3,5,6)

wilcox.test(teamX, teamY)
# pairwise.wilcox.test(c(teamX,teamY),rep(c("X","Y"), each=6))
```


## Q.2:

The manager of a national park wants to see if pollution levels in lake are reduced by  banning the  boat rental. This is measured by the rate of pollution every 60 minutes (8am 22pm: total of 15  measurements) in a day when rental is allowed, and in a day  wh en it is not. Here  are  the values  of water pollution:

With rental:  214, 159, 169, 202, 103, 119, 200, 109, 132, 142, 194, 104, 219, 119, 234
Without rental: 159, 135, 141, 101, 102, 168, 62, 167, 174, 159, 66, 118, 181, 171, 112
( ?? = 0.05 )
```{r}
with.rental <- c(214, 159, 169, 202, 103, 119, 200, 109, 132, 142, 194, 104, 219, 119, 234)
without.rental <- c(159, 135, 141, 101, 102, 168, 62, 167, 174, 159, 66, 118, 181, 171, 113)
```

A) Are these  two groups of sampling paired or independent? Why?
```{r}
# Paired: Same lake. (Time series sample)
```

B) Use  Wilcoxon signed rank test to calculate p - value. Based on the mentioned  significant level (?? = 0.05), what  do  you conclude about  banning the boat rental for  any improvements in terms of pollution of  the lake?
```{r}
#pairwise.wilcox.test(c(with.rental, without.rental), rep(c("with","without"),each=length(with.rental)), paired=TRUE)
wilcox.test(with.rental, without.rental, paired=TRUE)

#I would say, generally, that you would need more data to conclude that banning the boat rentals won't make a reduction in lake poillution. However, based exclusively on this test there isn't enough support at the a=0.05 confidence level. 
```

C) C alculate the sum of ranks assigned to the differences with positive sign ,  and the  sum of ranks assigned to the differences with negative sign: 
  1.Calculate the vector containing the differences (diff)
  2. Delete all differences equal to zero
  3. Check  the ranks of the differences, taken in absolute ,  recalling the signs of the  values of the differences (Calculate  diff.rank vector)
  4. C heck the sign to the ranks, recalling the signs of the values of the differences (Calculate  diff.rank.sign vector) 
  5. Calculate the sum of ranks assigned to the differences as a positive ( ranks.pos ) .  
  6. Calculate the sum of ranks assigned to the differences as a negative ( ranks. neg) 

```{r}
#1.
diff <- with.rental - without.rental
#2.
diff <- diff[diff != 0]
#3. 
diff.rank <- rank(abs(diff))
#4. 
diff.rank.sign <- sign(diff)
#5. 
sum(diff[diff.rank.sign > 0])
#6.
sum(diff[diff.rank.sign < 0])
```

  
## Q.3 :
Suppose we are going to compare two software  Zip and  Tar on archiving 20 different files. The  execution time (time from the beginning till the end of the  archiving files in milliseconds ) is  recorded as follows:

Zip execution time:  10, 44, 65, 77, 43, 44, 22, 66, 50, 100, 55, 99, 44, 23, 100, 88, 200, 220, 110, 551 
Tar execution time:  20, 55, 75, 60, 55, 88, 35, 33, 35, 80, 65, 82 , 47, 35, 97, 110, 250, 190, 111 , 600

```{r}
zip.time <- c(10, 44, 65, 77, 43, 44, 22, 66, 50, 100, 55, 99, 44, 23, 100, 88, 200, 220, 110, 551)
tar.time <- c(20, 55, 75, 60, 55, 88, 35, 33, 35, 80, 65, 82 , 47, 35, 97, 110, 250, 190, 111 , 600)

```

Your task is to find out if there is any significant difference in the execution times of two  different  archiving program s based on their running time  on same file.  

A) Are these two groups of sampling paired or independent? Why?

```{r}
#Paired; same file.
```

B)  Plot the distribution of these two groups of sampling  to see if they  have normal  distribution or not.
```{r}
plot(tar.time)
plot(zip.time)
# not very normal looking, bro
```



C )  Use  Wilcoxon signed rank test to calculate p - value. Based on the mentioned  significant level (?? = 0.05), what do you conclude about differences between the mean  of run - time of two different archiving programs?

```{r}
wilcox.test(zip.time, tar.time, paired=T)
# Insufficient evidence to reject H0.
```

