---
title: "CMTH642 Lab02"
author: "Geoffrey Clark"
date: "January 23, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1
We want to test with 95 percent confidence interval whether the volume of a shipment of lumber  is as usual (mu=39000 cubic feet). Use data< - rnorm(n, mean = , sd = ) to generate 75 shipments  with  mean:36500  and  sd:2000.  Use  set.seed(0)  before  rnorm  to  regenerate  the  same  data  if  required. On the simulated data test Ho: mu = 39000 Hint: t.test(data, mu = 39000)
```{r}
# a <- 95%
#H0: mu = 39000
set.seed(0)
data <- rnorm(n=75, mean=36500, sd=2000)
t.test(data, mu=39000)
```

## Q2

The results obtained for an intelligence test in 10 subjects are:

65, 78, 88, 55, 48, 95, 66, 57, 79, 81
i - Calculate the sample mean and standard deviation
ii - Use  a  one - sample  t - test  to  determine  whether  the  average  result  of  the  population  which  received the same test is equal to 75 using a significance level of 0.05.
```{r}
s_q2 <- c(65,78,88,55,48,95,66,57,79,81)
#i
sd(s_q2)
mean(s_q2)
#ii
t.test(s_q2, mu=75)
```

## Q3

A  bottle  filling  machine  is  set  to  fill  bottles  with  soft  drink  to  a  volume  of  500  ml.  The  actual  volume  is  known  to  follow  a  normal  distribution.  The  manufacturer  believes  the  machine  is  under - filling bottles. A sample of 20 bottles is taken and the volume of liquid inside is measured. 
The volumes were:
484.11, 459.49, 471.38, 512.01, 494.48, 528.63, 493.64, 485.03, 473.88, 501.59, 502.85, 538.08, 465.68, 495.03, 475.32, 529.41, 518.13, 464.32, 449.08, 489.27

i - Calculate the sample mean and standard deviation
ii - Use a one - sample  t - test to determine whether the bottles are being consistently under filled  using a significance level of 0.01
```{r}
#i
s_q3 <- c(484.11,459.49,471.38,512.01,494.48,528.63,493.64,485.03,473.88,501.59,502.85,538.08,465.68,495.03,475.32,529.41,518.13,464.32,449.08,489.27)

mean(s_q3)
sd(s_q3)

#ii
t.test(s_q3, mu=500, alternative="less", conf.level=0.99)


```

## Q4

An  outbreak  of  Salmonella - related  illness  was  attributed  to  ice  cream  produced  at  a  certain 
factory. Scientists measured the level of
Salmonella in 9 randomly sampled batches of ice cream. 
The levels (in MPN/g) were:
0.593 0.142 0.329 0.691 0.231 0.7930.5190.392 0.418
Is there evidence that the mean level of Salmonella in the ice cream is greater than 0.3 MPN/g?

```{r}
s_q4 <- c(0.593, 0.142, 0.329, 0.691, 0.231, 0.793, 0.519, 0.392, 0.418)

t.test(s_q4, alternative="greater", mu=0.3)

# P < 0.05 .: We can reject H0: mu = 0.3 
```

## Q5

Assuming  tha t  the  data  in  mtcars  follows  the  normal  distribution,  find  the  95%  confidence  interval  estimate  of  the  difference  between  the  mean  gas  mileage  of  manual  and  automatic  transmissions . (Hint: two sample t - test)

```{r}
d0 <- mean(mtcars[mtcars$am,]) - mtcars[!mtcars$am,]
```

## Q6

Consider the gain in weight of 19 fem ale rats between 28 and 84 days after birth. 12 were fed on  a high protein diet and 7 on a low protein diet. Using the following data, test the hypothesis that  there  is  no  difference  in  weight  gain  between  female  rats raised  on  a  high - protein  diet versus  t hose  raised  on  a  low - protein diet. Use a significance level of α = 0.05 and assume equal  var iances. ("Hint: var.equal=TRUE")
High protein: 134,146,104,119,124,161,107,83,113,129,97,12
Low protein: 70,118,101,85,107,132,94
```{r}
hp <- c(134,146,104,119,124,161,107,83,113,129,97,12)
lp <- c(70,118,101,85,107,132,94)

t.test(hp, lp) #Fail to reject H0: p-value = 0.4766
```

## Q7

Load the "MASS" package. In the immer dataset of the "MASS" library: we have: Y1 Yield in 1931,  Y2 Yield in 1932. Assuming that the data in immer follows the normal distribution, find the 95%  confidence  interval  estimate  of  the  difference  between  the  mean  barley  yields  between  years  1931 and 1932 (Hint: paired t - test). Get "p.value" in a variable pvalue and "statistic" in a variable  st. (Hint : ttest< - t.test (...,...,...) and then  names (ttest))

```{r}
#library(MASS)
test <- t.test(immer$Y1, immer$Y2)
statistic <- test$p.value
```

## Q8

```{r}
students <- c(25, 32, 18, 20)
chisq.test(students)
```

