---
title: "CMTH642 Lab06"
author: "Geoffrey Clark"
date: "February 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Q1:  the  a irquality dataset contains  the daily air quality measurements in New York,  from  May  to September 197 3 . Without assuming the data to have normal distribution, test at .05  significance level if the monthly ozone density in New York has identical data distributions from  May to September 1973. W hat  is null hypothesis?  W hat  is alternative hypothesis?

```{r}
# months <- split(airquality, airquality$Month)
# H0: All months have identical distributions
# H1: At least two months' distributions differ
kruskal.test(Ozone ~ Month, data=airquality)

```


Q2:  the test results for  English and Math tests for a class are given in table below:

English 56 75 45 71 62 64 58 80 76 61
Math 66 70 40 60 65 56 59 77 67 63

W ithout  any assumptions on the  distribution of the data,  investigate if math and English scores  are  correlated . A lso  investigate  if math and Eng l ish scores  have  any  monotonic relation.

```{r}
english <- c(56,75,45,71,62,64,58,80,76,61)
math <- c(66,70,40,60,65,56,59,77,67,63)
cor(english,math,method="spearman")
plot(english, math)
# Correlation, no monotonic relation
```

Q3: The table below shows the growth of an insect fe d upon a variety of sugar diets where e ach  column  represents  a  diet. C heck  if there is any significant evidence that shows diet affects the  growth.

  C  G  F  FG S  test
1 75 57 58 58 62 63
2 67 58 61 59 66 64
3 70 60 56 58 65 66
4 75 59 58 61 63 65
5 65 62 57 57 64 67
6 71 60 56 56 62 68
7 67 60 61 58 65 64
8 67 57 60 57 65 NA
9 76 59 57 57 62 NA
10 68 61 58 59 67 NA

```{r}
q3.c <- c(75,67,70,75,65,71,67,67,76,68)
q3.g <- c(57,58,60,59,62,60,60,57,59,61)
q3.f <- c(58,61,56,58,57,56,61,60,57,58)
q3.fg <- c(58,59,58,61,57,56,58,57,57,59)
q3.s <- c(62,66,65,63,64,62,65,65,62,67)
q3.test <- c(63,64,66,65,67,68,64,NA,NA,NA)
# q3.df <- data.frame(groups=c(q3.c,q3.g,q3.f,q3.fg,q3.s,q3.test), labels=as.factor(rep(1:6, each=10)))

q3.df <- data.frame(groups=c(q3.c,q3.g,q3.f,q3.fg,q3.s,q3.test), labels=as.factor(rep(1:6, each=10)), subjects=as.factor(1:10))

friedman.test(groups ~ labels|subjects, data=q3.df) #error
q3.df.rm.na <- q3.df[as.numeric(q3.df$subjects) < 8,]
friedman.test(groups ~ labels|subjects, data=q3.df.rm.na)

# I somehow got the above to work by splitting everything up into separate vectors
# (using the first 5 sets of values; does not work with NAs apparently)
# note: this apparently isn't a proper use of this test, I was more interested in figuring out
# why the function wouldn't run

vals <- c(q3.c,q3.g,q3.f,q3.fg,q3.s)
tmts <- rep(1:5, each=10)
blks <- rep(1:10,5)
friedman.test(vals ~ tmts | blks)

# Solutions use kruskal-wallis...

q3.df <- data.frame(C=q3.c, G=q3.g, F=q3.f, FG=q3.fg, S=q3.s, test=q3.test)

kruskal.test(q3.df)
boxplot(q3.df)
```


Q4: T able  below shows the  results for  survey of amphibians on 5 month of every year from 2004 to  2006 . The first column (count) represents the number of individuals captured.  W e wish to know  if there is any significant difference  on count  due to year. W hat  is null hypoth esis , what is  alternate hypothesis? 

count month year
1 2 1 2004
2 48 1 2005
3 40 1 2006
4 3 2 2004
5 120 2 2005
6 81 2 2006
7 2 3 2004
8 16 3 2005
9 36 3 2006
10 7 4 2004
11 21 4 2005
12 17 4 2006
13 2 5 2004
14 14 5 2005
15 17 5 2006
```{r}
#H0 : There is no difference between block distributions
#H1: The distribution of at least two blocks differ

q4.df <- data.frame(count=c(2,48,40,3,120,81,2,16,36,7,21,17,2,14,17), month=as.factor(rep(1:5, each=3)), year=as.factor(c(2004,2005,2006)))

friedman.test(count ~ month|year, data=q4.df)

```

