---
title: 'CMTH 642 LAB 10 '
output: pdf_document
---

Preparation: 

install.packages("class")

install.packages("gmodels")

library(class)

library(gmodels)

We will detect prostate cancer in this lab session. Download Prostate_Cancer.csv from the course shell. The data set contains patients who have been diagnosed with either Malignant (M) or Benign (B) cancer.

Questions:

1. Check data characteristics. Is there missing data? Prepare the data for analysis.
```{r}
pc <- read.csv("Prostate_Cancer.csv")
str(pc)
sum(is.na(pc))

```

2. Normalize the data set.
```{r}
pc$id <- NULL
# pc.norm <- pc[]
levels(pc$diagnosis_result) <- c(1,2) # B,M
```

3. Divide the data to training and testing groups.
```{r}
pc_train.index <- sample(1:nrow(pc), 0.7*nrow(pc))
pc_train <- pc[pc_train.index,]
pc_test <- pc[-pc_train.index,]
```

4. Use the KNN algorithm to predict the cancer.
```{r}
# pc_train_labels <- pc_train$diagnosis_result
# pc_test_labels <- pc_test$diagnosis_result
pc.knn <- knn(train=pc_train, test=pc_test, cl=pc_train$diagnosis_result, k=3)
```

5. Evaluate the model performance.
```{r}
CrossTable(y=pc_test$diagnosis_result, x=pc.knn, prop.chisq=F)
```

