---
title: "CMTH642 Assignment 3"
author: "Geoffrey Clark"
date: "March 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Prep

Preparation:  The  dataset is related to white Portuguese "Vinh o Verde" wine. For more info: https://archive.ics.uci.edu/ml/datasets/Wine+Quality

Import to R the following  file: http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv


```{r}
# wine.df <- read.csv("http://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv", header=T, sep=";")
wine.df <- read.csv("winequality-white.csv", header=T, sep=";")

# For Visualizing Correlation:
# install.packages("corrplot")
# library(corrplot)
```

QUESTIONS

1. Check data characteristics. Is there missing data? 
```{r}
sum(is.na(wine.df)) #0
str(wine.df)
sapply(wine.df, function(x) sum(is.na(x))) # No NAs
```

2. What is the correlation between the attributes other than wine quality? 
```{r}
(wine.cor <- cor(wine.df[,-12]))
corrplot(wine.cor, method="number") # I did this because I appreciate the visual

# highest correlation:
# density & residual.sugar: 0.84
# density & alcohol: -0.78

# I found a cool SO thread on how to do this programmatically:
# https://stackoverflow.com/questions/7074246/show-correlations-as-an-ordered-list-not-as-a-large-matrix#7074856

# as.data.frame(as.table(wine.cor))

# install.packages("reshape")
# library(reshape)

wine.cor.list <- wine.cor
wine.cor.list[wine.cor.list == 1] <- NA
wine.cor.list <- na.omit(melt(wine.cor.list))
wine.cor.list[order(-abs(wine.cor.list$value)),]
```


3. Graph the frequency distribution of wine quality .
```{r}
hist(wine.df$quality, freq=T, breaks=seq(2.5,9.5,1))
```


4. Reduce the levels of rating for quality to three levels as high, medium and low.
```{r}
# (max(wine.df$quality) - min(wine.df$quality))/3 # size of each level
# wine.quality.factors <- cut(wine.df$quality, c(2,6,10))
# sum(is.na(wine.quality.factors))
# wine.df$quality <- cut(wine.df$quality, c(2,4,7,11))
wine.df$quality <- cut(wine.df$quality, 3, labels=c("Low","Medium","High"))
```


5. Normalize the data set.
```{r}
levels(wine.df$quality) <- c(1, 2, 3) # kNN in Q7 won't work well with strings..
# scale() returns a matrix. Also requires numerical values (exclude class attribute (factor): column 12)
wine.df.norm <- as.data.frame(scale(wine.df[,-12])) 
wine.df.norm <- cbind(wine.df.norm, wine.df[,12]) # Re-add class attribute
names(wine.df.norm)[12] <- names(wine.df)[12] # Re-name the class attribute
```

6. Divide the data to training and testing groups.
```{r}
wine.train_index <- sample(1:nrow(wine.df.norm), 0.7*nrow(wine.df.norm))
wine.train <- wine.df.norm[wine.train_index,]
wine.test <- wine.df.norm[-wine.train_index,]
(nrow(wine.df) == (nrow(wine.train) + nrow(wine.test))) # Did we include all observations?
```

7. Use the KNN algorithm to predict the quality of wine using its attributes.
```{r}
# library(class) # needed for knn()
wine.knn <- list(k3=factor(),k5=factor(),k7=factor())
wine.knn$k3 <- knn(train=wine.train, test=wine.test, cl=wine.train$quality, k=3)
wine.knn$k5 <- knn(train=wine.train, test=wine.test, cl=wine.train$quality, k=5)
wine.knn$k7 <- knn(train=wine.train, test=wine.test, cl=wine.train$quality, k=7)
```


8. Evaluate the model performance .
```{r}
# library(gmodels) # Necessary for CrossTable()
# low: 1, medium: 2, high:3
# I prefer actual results along the top of the table: I find it easier to read
# This seems pretty standard, and is likewise on Wikipedia: https://en.wikipedia.org/wiki/Confusion_matrix
CrossTable(x=wine.knn$k3, y=wine.test$quality, prop.chisq = F)
CrossTable(x=wine.knn$k5, y=wine.test$quality, prop.chisq = F)
CrossTable(x=wine.knn$k7, y=wine.test$quality, prop.chisq = F)

# Different values of k produce very similar results with the best being a marginal k=5
# This classifier performs very well. However there is a noticable improvement after normalizing 
# the data with scale(). 
# 
# WITHOUT NORMALIZE:
# TP rates (k=3): 53.1%, 79.6% and 15.9% for Low, medium & high respectively

# WITH NORMALIZE:
# TP rates (k=3): 92.1%, 89.5%, 88.5%  
```



Save your R codes in an RMD file.
Send your  RMD and PDF  files to the course shell.